\propertitle

### 2019-06-18 maxNt_v5.tex
update

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-06-18T15:05:04+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal network\\ from subnetwork data%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

\newcommand{\appropto}{\mathrel{\vcenter{
  \offinterlineskip\halign{\hfil$##$\cr
    \propto\cr\noalign{\kern2pt}\sim\cr\noalign{\kern-2pt}}}}}
% https://tex.stackexchange.com/a/33547/97039
%\newcommand*{\fm}{f}
\newcommand*{\px}{p}

@@

The context we consider is as follows. We have recorded the spiking
activities of $n$ neurons for a given amount of time. Let's call these
neurons the \enquote{sample} or the \enquote{subnetwork}. Their spikes are
binned into $T$ time bins and binarized to $\set{0,1}$ values in each bin.
Denote by $\ya_{t}$ the number of neurons that fire during time bin $t$,
divided by the total number of neurons $n$; we call this the
\emph{normalized total activity} of the sample, or just its \enquote{total
  activity} or \enquote{activity} for short. Obviously
$\ya_{t} \in\set{0,1/n,\dotsc,(n-1)/n,1}$; if $\ya_{t}=0$, no neuron spiked
during time bin $t$; if $\ya_{t}=1$, all spiked at some point during time
bin $t$. For brevity, let's say \enquote{at $t$} for \enquote{during time
  bin $t$}.

\mynote{maybe move this paragraph to the intro?}Present-day technologies
enable the recording of neuronal activity from small brain regions
\mynote{be more specific}. The neurons that are recorded within the region
are not, and cannot yet be, specifically chosen from among the rest, owing
to several limiting factors -- for example, limitations in the precision
with which electrodes are inserted, or with which neurons are targeted by
viruses, and so on. In fact, the set of recorded neurons may even change
during very long recordings or across experimental trials. We assume that
there's a region, comprising a network of $N$ neurons, from which other
sets of neurons had the same probability of being recorded as the set of
$n$ neurons that has actually been recorded. We call this the \enquote{full
  network}\mynote{how about calling it \enquote{the pool}?}. During the
recording, the normalized total activity of these $N$ neurons at $t$ was
$\yA_{t}$. We don't know the values $\yA_{t}$ at each $t$; we only know for
certain that $\yA_{t} \in \set{0,1/N,\dotsc,(N-1)/N,1}$ and that
$N\yA_{t} \ge n\ya_{t}$ for obvious reasons. For the moment we assume that
we know $N$; in \sect*** we discuss the consequences of our lack of precise
knowledge about this number.

Our questions concern general features of the total activity $\yA$ of the
full network during and after the recording. For example: what was its
frequency distribution in time during the recording? How much total
activity should we expect at a time bin during the recording? And at a new
time bin? We cannot answer these questions with certainty; we can only give
distributions of probability over their possible answers. Our approach
gives such probability distributions.

@@

The idea behind the approach is easily summarized:
\begin{enumerate}[label=(\alph*)]
  \item\label{item:maxent_step} we build a distribution for the total
  activity of the full network using the maximum-entropy method;
  \item\label{item:sample_step} the constrained averages used in the
  maximum-entropy method for the full network are, in turn, determined via
  sampling theory from the constrained averages for the sample.
\end{enumerate}
Let's discuss these points in detail.

Regarding~\ref{item:maxent_step}, we assume that you're familiar with the
maximum-entropy method. In short, it amounts to a pair of prescriptions:
choose the distribution having maximum Shannon entropy among those
satisfying specific convex constraints, such as fixed averages; and judge
these averages to be equal to measured averages. We add two remarks about
this method that are seldom made in the literature. First, the
\enquote{distribution} given by this method, which we denote by $\px(\yA)$,
is the zeroth-order approximation of four different distributions for the
full network:
\begin{itemize}
\item the most probable \emph{frequency} distribution for the total
  activity across the \emph{recorded} bins, %$\yFF(\yA)$,
\item the most probable \emph{frequency} distribution for the total
  activity in a very long run of \emph{new} time bins, $\ynuu(\yA)$,
\item the \emph{probability} distribution for the value of the total
  activity at any time bin among those \emph{recorded},
  %$\yp(\yA_{t} \| \yH)$ with $1\le t \le T$,
\item the \emph{probability} distribution for the value of the total
  activity at a \emph{new} time bin.
  % $\yp(\yA_{t} \| \yH)$ with $t<1$ or $t>T$
\end{itemize}
The maximum-entropy distribution is thus an approximate answer to four
completely different inferences. Note that although the four distributions
above numerically coincide in this approximation, they differ in
higher-order approximations. We discuss their differences in \sect***.
Second, the maximum-entropy method based on the Shannon entropy implicitly
makes specific assumptions about the probabilities for the long-run
frequency distributions
\citep{jaynes1986d_r1996,portamana2009,portamana2017}. We discuss these
assumptions in \sect***.

@@

We can therefore apply the maximum-entropy method to obtain a distribution
for the full network, by constraining its $m$th factorial moment to be
equal to the sample's recorded average of
$\tbinom{n\ya}{m}\;\tbinom{n}{m}^{-1}$, for as many $m$ as we please with
$1\le m \le n$. The constraints on the distribution $\px(\yA)$ for
the full network are
\begin{equation}
  \label{eq:final_constraint}
  \frac{1}{T}\sum_{t}\binom{n\ya_{t}}{m}\,\binom{n}{m}^{-1}
  =
   \sum_{\yA}
  \binom{N \yA}{m}\,  \binom{N}{m}^{-1}\, \pf(\yA)
\end{equation}
for any $m$s we wish.

>>>>

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-06-18T16:16:26+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal network\\ from subnetwork data%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

\newcommand{\appropto}{\mathrel{\vcenter{
  \offinterlineskip\halign{\hfil$##$\cr
    \propto\cr\noalign{\kern2pt}\sim\cr\noalign{\kern-2pt}}}}}
% https://tex.stackexchange.com/a/33547/97039
%\newcommand*{\fm}{f}
\newcommand*{\px}{p_{\text{me}}}

@@

The context we consider is as follows. We have recorded the spiking
activities of $n$ neurons for a given amount of time. Let's call these
neurons the \enquote{sample} or the \enquote{subnetwork}. Their spikes are
binned into $T$ time bins and binarized to $\set{0,1}$ values in each bin.
Denote by $\ya_{t}$ the number of neurons that fire during time bin $t$,
divided by the total number of neurons $n$; we call this the
\emph{normalized total activity} of the sample, or just its \enquote{total
  activity} or \enquote{activity} for short. Obviously
$\ya_{t} \in\set{0,1/n,\dotsc,(n-1)/n,1}$; if $\ya_{t}=0$, no neuron spiked
during time bin $t$; if $\ya_{t}=1$, all spiked at some point during time
bin $t$. For brevity, let's say \enquote{at $t$} for \enquote{during time
  bin $t$}. From the activities $\set{\ya_{t}}$ we can count how many times
the activity $\ya=0$, $\ya=1/n$, and so on each appeared during the
recording, obtaining the measured relative-frequency distribution
$\yff(\ya) \defs \yf$.

\mynote{maybe move this paragraph to the intro?}Present-day technologies
enable the recording of neuronal activity from small brain regions
\mynote{be more specific}. The neurons that are recorded within the region
are not, and cannot yet be, specifically chosen from among the rest, owing
to several limiting factors -- for example, limitations in the precision
with which electrodes are inserted, or with which neurons are targeted by
viruses, and so on. In fact, the set of recorded neurons may even change
during very long recordings or across experimental trials. We assume that
there's a region, comprising a network of $N$ neurons, from which other
sets of neurons had the same probability of being recorded as the set of
$n$ neurons that has actually been recorded. We call this the \enquote{full
  network}\mynote{how about calling it \enquote{the pool}?}. During the
recording, the normalized total activity of these $N$ neurons at $t$ was
$\yA_{t}$; the relative frequencies of the various activity levels were
$\yFF(\yA) \defs \yF$. We don't know the values $\yA_{t}$ at each $t$ or
the frequency distribution $\yF$. We only know for certain that
$\yA_{t} \in \set{0,1/N,\dotsc,(N-1)/N,1}$ and that $N\yA_{t} \ge n\ya_{t}$
for obvious reasons. For the moment we assume that we know $N$; in \sect***
we discuss the consequences of our lack of precise knowledge about this
number.

Our questions concern general features of the total activity $\yA$ of the
full network during and after the recording. For example: what was its
frequency distribution $\yF$ during the recording? How much total activity
should we expect at a time bin during the recording? And at a new time bin?
We cannot answer these questions with certainty; we can only give
distributions of probability over their possible answers. Our approach
gives such probability distributions.

@@

The idea behind the approach is easily summarized:
\begin{enumerate}[label=(\alph*)]
  \item\label{item:maxent_step} we build a distribution $\px(\yA)$ for the total
  activity of the full network using the maximum-entropy method;
  \item\label{item:sample_step} the constrained averages used in the
  maximum-entropy method for the full network are, in turn, determined via
  sampling theory from the constrained averages for the sample.
\end{enumerate}
Let's discuss these points in detail.

Regarding~\ref{item:maxent_step}, we assume that you're familiar with the
maximum-entropy method. In short, it amounts to a pair of prescriptions:
choose the distribution having maximum Shannon entropy among those
satisfying specific convex constraints, such as fixed averages; and judge
these averages to be equal to measured averages. We add two remarks about
this method that are seldom made in the literature. First, the distribution
$\px(\yA)$ given by this method is the zeroth-order approximation of four
different distributions for the full network:
\begin{itemize}
\item the most probable \emph{frequency} distribution for the total
  activity across the \emph{recorded} bins, %$\yFF(\yA)$,
\item the most probable \emph{frequency} distribution for the total
  activity in a very long run of \emph{new} time bins, $\ynuu(\yA)$,
\item the \emph{probability} distribution for the value of the total
  activity at any time bin among those \emph{recorded},
  %$\yp(\yA_{t} \| \yH)$ with $1\le t \le T$,
\item the \emph{probability} distribution for the value of the total
  activity at a \emph{new} time bin.
  % $\yp(\yA_{t} \| \yH)$ with $t<1$ or $t>T$
\end{itemize}
The maximum-entropy distribution is thus an approximate answer to four
completely different inferences. Note that although the four distributions
above numerically coincide in this approximation, they differ in
higher-order approximations. We discuss their differences in \sect***.
Second, the maximum-entropy method based on the Shannon entropy implicitly
makes specific assumptions about the probabilities for the long-run
frequency distributions
\citep{jaynes1986d_r1996,portamana2009,portamana2017}. We discuss these
assumptions in \sect***.

@@

We can therefore apply the maximum-entropy method to obtain a distribution
for the full network, by constraining its $m$th factorial moment to be
equal to the sample's recorded average of
$\tbinom{n\ya}{m}\;\tbinom{n}{m}^{-1}$, for as many $m$ as we please with
$1\le m \le n$. The constraints on the distribution $\px(\yA)$ for
the full network are
\begin{equation}
  \label{eq:final_constraint}
  \frac{1}{T}\sum_{t}\binom{n\ya_{t}}{m}\,\binom{n}{m}^{-1}
  =
   \sum_{\yA}
  \binom{N \yA}{m}\,  \binom{N}{m}^{-1}\, \px(\yA)
\end{equation}
for any $m$s we wish.
